\chapter{Gaps}\label{chap:gaps}

This chapter collects some gaps, but not all and sorts them into categories. 

\section{Gaps in graphics}\label{sec:gapGraph}

Only figures created with \tool{xfig} and stored as files PDF and PTX 
may be integrated into a \LaTeX{} document. 
This could be extended to a broader variety of export file formats. 
The problem is, that fig-files to not contain information on the export
format. 
This has to be either given elsewhere in a config file 
or determined by pre-parsing the TEX files. 

There is no support for pictures in \gls{gif}-format 
but maybe a converter to \gls{png} is all needed. 

\section{Build mechanism}\label{sec:gapBuild}

There is no proper make-mechanism taking dependencies into account. 
Thus, all documents in all formats specified are remade, 
whether they changed or not. 

Also, if more than one target is created from one \LaTeX{} source, 
common steps are redone for each target. 
E.g.~if PDF and HTML are created, 
PDF creation is done twice and if PDF, HTML, ODT and DOCX are created, 
ODT is done twice (once for ODT second for DOCX) 
and PDF is done even trice: 
once for pv itself, once for ODT and once for DOCX. 

\section{Indices}\label{sec:gapIndices}

% This has to be removed. 
Creating more than one index is supported only 
via package \pkg{splitidx} in conjunction with \texttt{SplitIndex}. 
There are the following packages also supporting multiple indices 
but not supported officially: 
\pkg{index} described in~\cite{IndexP}, 
\pkg{amsmidx} described in~\cite{AmsmidxP} and 
\pkg{imakeidx} described in~\cite{ImakeidxP}.
Note that the package \pkg{multind} is obsolete. 

\section{Glossaries}\label{sec:gapGlossaries}

According to~\cite{GloP4_54}, Section~1.3, 
there are various options to create a glossary, 
whereas this software supports option two only described in Section~1.3.2, 
which uses \texttt{makeindex} for indexing. 
Currently, indexing with \texttt{xindy} is not supported. 
The last two options are available only with package \pkg{glossaries-extra} 
which this software will support in later versions. 
% planned in 2.2. 
% Then this technique is devoted a separate section and this becomes the preferred usage. 

By default, package \pkg{glossaries} creates a single ``main glossary'', 
which can be switched off specifying the option \texttt{nomain} 
described in Section~2.6. 
In this case at least, more specific glossary types must be specified. 
This can be done by options like \texttt{acronyms} described in Section~2.7 
or the \texttt{symbols},
\texttt{numbers} or \texttt{index} options described in Section~2.9. 
As the \texttt{index} option collides with indexing as performed by this software, 
the option \texttt{index} is not allowed. 

The package \pkg{glossaries} itself 
supports new glossary types via the command
%
\begin{lstlisting}[language=TeX, basicstyle=\small]
\newglossary[log-ext]{name}{in-ext}{out-ext}{title}[counter]
\end{lstlisting}
%
described in~\cite{GloP4_54}, Section~9. 
In fact, the glossary types accessible via options and even the main glossary 
are defined internally that way. 

Although the glossary algorithm of this software, 
in particular rerun management as described in Section~\ref{sec:runRerunAux}
can create any kind of glossaries created with \cmd{newglossary}, 
and it can also clean up files created in conjunction with glossaries 
as long as the file endings do not contain ``\texttt{.}'',
defining new glossary types is not recommended 
because \texttt{latexmk} cannot mimic this 
with a fixed \texttt{.latexmkrc} file, neither in creation rules 
nor in patter for files for deletion  
and because collision, e.g. with indexing, cannot be excluded. 

Reading~\cite{GloP4_54}, Section~13.1, the glossary option \texttt{index} 
seems to allow creating indices through the \pkg{glossaries} package 
making any index-package obsolete. 
This software does not support that technique offered by the package \pkg{glossaries}. 

For development given the \LaTeX{} main file \texttt{xxx.tex}, 
the files \texttt{xxx.pdf}, \texttt{xxx.pdf}, \texttt{xxx.synctex.gz} 
and \texttt{xxx.log} are vital. 
Thus, it would be fine to have a goal which touches these files 
or to have a parameter to touch these prior to creation 
to avoid that these are cleaned up after the run. 
This is an alternative to setting parameter \texttt{cleanup} to \texttt{false}. 
On the other hand, goal \texttt{grp} creating graphics 
in conjunction with a development tool like \tool{vscode}, 
allows to compile a \LaTeX{} main file in that tool 
and thus to access \texttt{xxx.log} and \texttt{xxx.pdf}. 

There are lots of possible improvements to be done on the goal \emph{check}. 


The ant-task does not allow creating single formats, e.g.~pdf selectively. 

The ant-build is not completed: tests are not run and 
test runs are no prerequisite for installation. 

This manual is not finished. 
To test the overall functionality of the maven-plugin and of the ant-task 
described here, this manual is created through plugin and task. 

Support for djvu via pdf2djvu: 
pdf2djvu -o output\_file input\_file

pdf2dsc
(ps with document structuring convention) 

pdf2svg is not so useful. 

pdftohtml -c is also not bad, 

consider also pdftocairo for creation of tiff and ps and many others. 
